---
title: 'Lab Week 5: time series and a map!'
author: "Gabriel De La Rosa"
date: "2/6/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Attach packages:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(paletteer)
library(tsibble)
library(fable)
library(fabletools)
library(feasts)
library(forecast)
library(sf)
library(tmap)
library(mapview)
library(naniar)
```


## Monthly US energy consumption (renewables)


```{r}
us_renew <- read_csv(here::here("data", "renewables_cons_prod.csv")) %>% 
  clean_names()
```

- Make description all lowecase
- only keep observations where the "description" variable contains "consumption"
- remove any observations where "description" variable contains "total"

```{r}
renew_clean <- us_renew %>% 
  mutate(description = str_to_lower(description)) %>% 
  filter(str_detect(description, pattern = "consumption")) %>% 
  filter(!str_detect(description, pattern = "total"))
```

# Convert `yyyymm` to date

# To work with feast/fable, we want to convert the date to year month using the tsibble package

```{r}
renew_date <- renew_clean %>% 
  mutate(yr_mo_day = lubridate::parse_date_time(yyyymm, "ym")) %>% 
  mutate(month_sep = yearmonth(yr_mo_day)) %>% 
  mutate(value = as.numeric(value)) %>% 
  drop_na(month_sep, value)


# Make a version where month and year are separate columns:

renew_parsed <- renew_date %>% 
  mutate(year = year(yr_mo_day)) %>% 
  mutate(month = month(yr_mo_day, label = TRUE))

```


# Let's look at it:

```{r}

renew_gg <- ggplot(data = renew_date, aes(x = month_sep, 
                                          y = value, 
                                          group = description)) +
  geom_line(aes(color = description)) 
  
renew_gg

```

Update colors with paletteer colors:

```{r}

renew_gg +
  scale_color_paletteer_d("palettetown::tyranitar")

```


### Corerce renew_parsed to a tsibble

key = main variable you're looking at (not required)
indes = tsibble compatible time series 

```{r}
renew_ts <- as_tsibble(renew_parsed, key = description, index = month_sep)
```

Let's look at our ts data in a few different ways:

```{r}

# Autoplot knows which value is your key!

renew_ts %>% autoplot(value)

# gg_subseries breaks up years and months for you

renew_ts %>% gg_subseries(value)

# We can also explore a season plot: within each season, how have things shifted over time

# Whoop. This shit don't work.

## renew_ts %>% gg_season(value, n = year))

# But, we can make this with ggplot!!!

ggplot(data = renew_parsed, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year)) +
  facet_wrap(~description,
             ncol = 1, 
             scales = "free",
             strip.position = "right")

```

### Let's just look at the hydroelectric consumption:


```{r}
hydro_ts <- renew_ts %>% 
  filter(description == "hydroelectric power consumption")


hydro_ts %>% autoplot(value)
hydro_ts %>% gg_subseries(value)

ggplot(data = hydro_ts, aes(x = month, y = value, group = year)) +
  geom_line(aes(color = year))

```

### What if I want the quarterly average consumption for hydro?

We can use a function `index_by()` in the tsibble package

```{r}
hydro_quarterly <- hydro_ts %>% 
  index_by(year_qu = ~(yearquarter(.))) %>% 
  summarize(avg_consumption = mean(value))
```


### Let's decompose that hydro_ts data and look at different components


STL decomposes by Loess smoothing

```{r}
dcmp <- hydro_ts %>% 
  model(STL(value ~ season(window = 5)))
  
components(dcmp) %>% autoplot()

```


```{r}
# Viewing residuals is easy too

hist(components(dcmp)$remainder)

```

Now, lets look at the autocorrelation

```{r}
hydro_ts %>% 
  ACF(value) %>% 
  autoplot()
```

### DANGER DANGER do a lot of reading before forecasting....


```{r}
hydro_model <- hydro_ts %>% 
  model(
    ARIMA(value)
  ) %>% 
  fabletools::forecast(h = "4 years")


hydro_model %>% autoplot(filter(hydro_ts, year(month_sep) > 2010))

```

### Make a world map!


mapview is awesome for a quick and dirty look at spatial data
```{r}

world <- read_sf(here::here("data", "TM_WORLD_BORDERS_SIMPL-0.3-1"),
                 layer = "TM_WORLD_BORDERS_SIMPL-0.3")

mapview(world)

```




